---
title: "Top 6 Most Difficult Sports"
echo: false
message: false
warning: false
---

There is a surprising difference in the difficulty for the top 6 "hardest" sports. The data in this project/plot were pulled from an online database on Kaggle with the different requirements for different sectors of skills required for a variety of sports. To make the graph, we mapped the data to find the top six hardest sports in terms of difficulty. Comparing the top six hardest sports in the category of difficulty, we see that boxing is the hardest sport.

```{r}
#| label: setup
#| echo: false
#| message: false

library(tidyverse)
library(primer.data)
library(ggplot2)
library(readr)
library(dplyr)
library(grid)
library(gridExtra)
```

```{r}
# Read the CSV file
toughest_sport <- read_csv("toughestsport.csv")

# Calculate the average score for each sport across all categories
toughest_sport <- toughest_sport %>%
  rowwise() %>%
  mutate(average_score = mean(c_across(Endurance:Durability), na.rm = TRUE)) %>%
  ungroup()

# Select top 6 sports based on a specific column
get_top_sports <- function(data, column) {
  data %>% arrange(desc(!!sym(column))) %>% slice(1:6)
}

# Define a manual color scale for the sports
sport_colors <- c(
  "Boxing" = "#1f78b4", "Ice Hockey" = "#33a02c", "Football" = "#e31a1c",
  "Basketball" = "#ff7f00", "Wrestling" = "#6a3d9a", "Martial Arts" = "#b15928",
  "Tennis" = "#a6cee3", "Gymnastics" = "#b2df8a", "Baseball" = "#fb9a99",
  "Soccer" = "#fdbf6f"
)

# Ensure all sports are represented in the legend by adding rows for missing sports with NA values
all_sports <- data.frame(SPORT = names(sport_colors))
toughest_sport <- full_join(toughest_sport, all_sports, by = "SPORT")

# Create a bar plot
create_bar_plot <- function(data, y_var, title, y_label) {
  ggplot(data, aes(x = reorder(SPORT, !!sym(y_var)), y = !!sym(y_var), fill = SPORT)) +
    geom_bar(stat = "identity") +
    coord_flip() +
    labs(title = title, x = "Sport", y = y_label) +
    theme_minimal() +
    scale_fill_manual(values = sport_colors, drop = FALSE) +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold"),  # Center and bold title
      axis.title.x = element_text(face = "bold"),             # Bold x axis title
      axis.title.y = element_text(face = "bold")              # Bold y axis title
    )
}

# Get top 6 sports by overall difficulty
top6_difficulty <- get_top_sports(toughest_sport, "average_score")

# Create bar plot for overall difficulty
p2 <- create_bar_plot(top6_difficulty, "average_score", "Top 6 Sports by Overall Difficulty", "Average Score")

# Print the plot
print(p2)
```

